### 2017-05-10 01-intro.tex
Begin working on picture refactor

Starting from an existing implementation in Python/Numpy, we will present
how this algorithm can be implemented much more efficiently in C++. Then,
we will describe our two key contributions: first, how to scale up our
implementation to run on GPUs and multicore CPUs, second, to scale out
our implementation to overcome the memory bounds posed by GPUs or general
single-machine systems.
% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem. In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.

@@

The second part of our work, scaling out to the largest publicly available
community graphs, also necessitated overcoming several challenges. First,
the algorithm's state grows rapidly with larger graphs and number of latent
communities.  Since the full state for the larger graphs does not fit in a
single machine's memory, it is partitioned and distributed across a cluster
of machines. Second, to access the full state, each cluster node must read
remote memory hosted by its peers. We leveraged RDMA to limit the high latency
of such operations and increase the communication bandwidth. To reduce the
latency further, we pipelined the algorithm's computations such that data can
be fetched in advance over the network. Finally, the algorithm's computation
is effectively distributed across the cluster nodes and parallelized further
within each node by exploiting their multi-core CPUs.

>>>>

Starting from an existing implementation in Python/Numpy, we will present
how this algorithm can be implemented much more efficiently in C++. Then,
we will describe our two key contributions: first, how to scale up our
implementation to run on GPUs and multicore CPUs, second, to scale out
our implementation to overcome the memory bounds posed by GPUs or
single-machine systems in general.
% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem. In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.

@@

The second part of our work, scaling out to the largest publicly available
community graphs, also necessitated overcoming several challenges.
The algorithm's state grows rapidly with larger graphs and number of latent
communities.  Since the full state for the larger graphs does not fit in a
single machine's memory, it is partitioned and distributed across a cluster
of machines. Hence, to access the full state, each cluster node must read
remote memory hosted by its peers. We leveraged RDMA to improve the latency and
bandwidth
of such operations. To reduce the
latency further, we pipelined the algorithm's computations by fetching
data in advance over the network.
\begin{comment} Finally, the algorithm's computation
is effectively distributed across the cluster nodes and parallelized further
within each node by exploiting their multi-core CPUs.
\end{comment}

### 2017-05-10 02-background.tex
Begin working on picture refactor

SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points. Then, Metropolis-Hastings accept-reject tests are applied. Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.

Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex. By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eq.~\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}. In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$. Then, the update rule of SGRLD is:
\bea
\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
\eea
here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.

>>>>

SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points. Then, Metropolis-Hastings accept-reject tests are applied. Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.

\looseness=-1
Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex. By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eq.~\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}. In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$. Then, the update rule of SGRLD is:
\bea
\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
\eea
here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.

### 2017-05-10 02-related.tex
Initial merge of intro, conclusions
Begin working on picture refactor

\section{Related Work}

\subsection{Related work on Accelerators for Bayesian methods}

In this section, we discuss related machine learning projects that use
accelerators.
Immediately when GPUs evolved into generally programmable accelerators, machine
learners recognized the potential of GPUs for their science.
Much research
has been devoted to deep-learning implementations on GPUs. However, Mixed
Membership deduction (MMSB) is usually considered untractable for deep-learning
algorithms. In contrast, our algorithm belongs to the family of Bayesian
(approximative, stochastic) methods, where efficient algorithms for large-scale
data have recently been developed. A broad survey of Bayesian methods is given
in~\cite{DBLP:journals/corr/Zhu0H14}. Broadly speaking, Bayesian approximative
algorithms are subdivided into variational and Monte Carlo methods; Markov
Chain Monte Carlo (MCMC) is a subgroup of the latter. GPUs have been put to
use in MCMC algorithms, e.g.\ \cite{journals/bioinformatics/MedlarGSBK13}
uses GPUs to analyse parental linkage patterns in a biology context and
\cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
Variational inference on MMSB is described
in~\cite{DBLP:conf/nips/GopalanMGFB12}, but they don't use GPUs. Latent
Dirichlet Allocation, another variety of Bayesian Approximation, is used on
GPUs by~\cite{DBLP:conf/nips/YanXQ09}.

@@

\subsection{Related work on parallel community detection}

Among previous work on parallelizing MMSB algorithms, we mention the Online
Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs.
There are a few projects that did a distributed implementation
for community detection on very large network graphs. In contrast to our work,
they all detect non-overlapping communities: \cite{Bu2013246} investigate a
large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
multi-core machines.

>>>>

\section{Related Work}
\label{sec-related}

\textbf{\textit{\large Need to rewrite to better fit the Background Section}}

Immediately when GPUs evolved into generally programmable accelerators, machine
learners recognized the potential of GPUs for their science.
Much research
has been devoted to deep-learning implementations on GPUs. However, Mixed
Membership deduction (MMSB) is usually considered untractable for deep-learning
algorithms. In contrast, our algorithm belongs to the family of Bayesian
(approximative, stochastic) methods, where efficient algorithms for large-scale
data have recently been developed. A broad survey of Bayesian methods is given
in~\cite{DBLP:journals/corr/Zhu0H14}. Broadly speaking, Bayesian approximative
algorithms are subdivided into variational and Monte Carlo methods; Markov
Chain Monte Carlo (MCMC) is a subgroup of the latter. GPUs have been put to
use in MCMC algorithms, e.g.\ \cite{journals/bioinformatics/MedlarGSBK13}
uses GPUs to analyse parental linkage patterns in a biology context and
\cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
Variational inference on MMSB is described
in~\cite{DBLP:conf/nips/GopalanMGFB12}, but they don't use GPUs. Latent
Dirichlet Allocation, another variety of Bayesian Approximation, is used on
GPUs by~\cite{DBLP:conf/nips/YanXQ09}.

@@

There are a few projects that did a distributed implementation
for community detection on very large network graphs. In contrast to our work,
they all detect non-overlapping communities: \cite{Bu2013246} investigate a
large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
multi-core machines.

### 2017-05-10 distr-03-design.tex
Some nitpicks
Begin working on picture refactor

\subsection{System Design}
\label{design-section}

@@

The SG-MCMC algorithm described in the previous section has an abundance of
opportunities for parallelism, for the multi-threaded shared-memory type as
well as for distributed-memory parallelism. The benefit of multi-threaded
parallelism is the speedup of the computation. Additionally, a distributed implementation
allows the storage of data in the collective memory of a cluster
and increases aggregate memory bandwidth. Both of these parameters scale with the number of
machines. The downside of a distributed implementation is that it requires
considerably more programming effort.

@@

This section describes how we created a distributed version of the C++
algorithm. We didn't use the advanced intra-node parallelisation techniques
described in the previous section, but found that straightforward
multi-threaded parallelization techniques sufficed: the distributed performance
is dominated by (remote) memory latency, not by node parallelism.
The details of the multi-threaded parallelism
will be discussed only where appropriate. The
distributed design follows a master-worker paradigm, where
the master controls the sequence of parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
MPI~\cite{Forum:1994:MMI:898758}.

>>>>

% \subsection{System Design}
\label{design-section}

@@

The SG-MCMC a-MMSB algorithm has an abundance of
opportunities for parallelism, for the multi-threaded shared-memory type as
well as for distributed-memory parallelism. The benefit of multi-threaded
parallelism is the speedup of the computation. Additionally, a distributed implementation
allows the storage of data in the collective memory of a cluster
and increases aggregate memory bandwidth. Both of these parameters scale with the number of
machines. The downside of a distributed implementation is that it requires
considerable implementation effort.

@@

This section describes how we created a distributed version of the C++
algorithm. We didn't use the advanced intra-node parallelisation techniques
described in the previous section, but found that straightforward
multi-threaded parallelization techniques sufficed: the distributed performance
is dominated by (remote) memory latency, not by node parallelism. As our
analysis will show, our pipelining optimizations will fully overlap the
parallel calculations with (pre)loading the data from remote hosts.
The details of the multi-threaded parallelism
will be discussed only where appropriate. The
distributed design follows a master-worker paradigm, where
the master controls the sequence of parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
MPI~\cite{Forum:1994:MMI:898758}.

### 2017-05-10 distr-04-evaluation.tex
Begin working on picture refactor

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

>>>>

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
  \caption{(a)~Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b)~Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

### 2017-05-10 gpu-05-03-accelerator.tex
Begin working on picture refactor

% CPU
\begin{figure}[htb]
\centering
\epsfig{file=plots/cpu-vector.eps, width=\columnwidth}
\caption{Performance of CPU for varying vector width.}
\label{FIG-CPU-VECTOR}
\end{figure}

@@

\subsubsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on Nvidia's GTX TitanX,
GTX980, K40c and K20m.

% TITANX
\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{titanx-w1-sweep}
\end{figure*}

% TITANX K=2K
\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=2048, M=4096, n=32.}
  \label{titanx-w1-sweep-2k}
\end{figure*}

Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX TitanX GPU
for all 10 strategy flavors, without explicitly vectorizing the kernels. The
x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
the total execution time of 1000 invocations of the \textit{update\_phi} kernel. The
naive and shared strategies are labeled NAIVE and SHARED respectively. Further,
each flavor of the code generation strategy is labeled by GEN followed by the 3
choices that identify it.
%
The wide performance range inhibits the readability of the figure,
therefore, a more focused Figure~\ref{titanx-w1-sweep}(b) is provided. As would
be expected, the naive strategy exhibits the worst performance over all
thread block
sizes, as it does not explicitly cache repeated device memory read operations.
The SHARED and \textit{GEN-SSS} strategies come next in terms of performance.
Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
differ in one aspect. Namely, \textit{GEN-SSS} explicitly unrolls the internal
loops of the kernel. However, there is no significant performance difference
between them. The other flavors of the code generation strategy attain higher
performance as they unroll internal loops as well as cache data in
registers. The TitanX obtains the best performance with the \textit{GEN-RSS}
strategy and a thread block size of 64. The results of explicitly vectorized kernels
are omitted as they obtain worse performance on the TitanX.

@@

% K40
\begin{figure*}[t]	% [hbt]
  \centering
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the K40c GPU, with
  explicit kernel vectorization of width 2, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{k40-w2-sweep}
\end{figure*}

% \looseness=-1
Figures~\ref{k40-w2-sweep} (a) and (b) present the performance of the K40c GPU
for the same experimental configuration as shown for the TitanX in
Figure~\ref{titanx-w1-sweep} with one exception. Namely, the code version used
for this plot is explicitly vectorized with a vector width of~2. The results
for the other code versions with vector width 4 and no vectorization are
omitted as they exhibit lower performance.
%
Surprisingly, Figure~\ref{k40-w2-sweep} shows that the NAIVE strategy
outperforms SHARED and some of the GEN strategies. This can be explained by the
unique properties of the Tesla Super Computing line of products to which the
K40c belongs. These GPUs include enhanced L2 caching mechanisms that
accelerate repeated and sparse memory accesses. This is especially advantageous
as it caches repeated reads across streaming multiprocessors. However, the
highest performance is attained by \textit{GEN-RRS} which explicitly employs
registers for both $Probs_i$ and $Grads_i$.

The performance results obtained from the GTX TitanX and Tesla K40c GPUs
reinforce the importance of customizing compute kernels to each GPU's specific
architecture and capabilities. For instance, each GPU achieved its highest
performance by employing a different strategy. Moreover, each GPU displayed
different strategy-performance orderings.

@@

\subsubsection{Comparison of Compute Devices}
Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
GTX TitanX, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
sequential C++ version. These results are consistent with the relative
capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
the TitanX achieves the highest speedup of 86 relative to the baseline. At the
other end, the parallel CPU version attains a speedup factor of~20.9.

As vectorization is a popular GPU optimization, it is important to note that
both GTX GPUs achieved their highest performance with non-vectorized kernels.
On the other hand, the Tesla GPUs obtained the best performance with a vector
width of~2. Figure~\ref{gpus-v-sweep} presents the execution time of the best
performing strategy for each of the 4 GPUs. In this figure, the performance is
normalized over the non-vectorized kernel version for each GPU. For instance,
the TitanX exhibits an overhead factor of approximately 1.8 when using a vector
width of~4. On the other hand, the K40c improves its performance by roughly
20\% when it uses a vector width of~2 compared to the non-vectorized kernel.
Therefore, explicit vectorization of the kernels can be either useful or
harmful depending on the GPU architecture and the specific problem it is
applied to.

>>>>

% CPU
\begin{figure}[tb]
\centering
\epsfig{file=plots/cpu-vector.eps, width=\columnwidth}
\caption{Performance of CPU for varying vector width.}
\label{FIG-CPU-VECTOR}
\end{figure}

@@

\subsubsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on Nvidia's GTX~\mbox{Titan-X},
GTX980, K40c and K20m.

% TITANX
\begin{figure*}[t]	% [htb]
  \centering
  \setlength\tabcolsep{0em}
  \begin{tabular}{c c c}
  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=0.33\textwidth} &
  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=0.33\textwidth} &
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.33\textwidth}
  \end{tabular}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the
  \mbox{Titan-X} GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{titanx-w1-sweep}
\end{figure*}

\begin{comment}
% TITANX K=2K
\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the
  \mbox{Titan-X} GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=2048, M=4096, n=32.}
  \label{titanx-w1-sweep-2k}
\end{figure*}
\end{comment}

Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX~\mbox{Titan-X} GPU
for all 10 strategy flavors, without explicitly vectorizing the kernels. The
x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
the total execution time of 1000 invocations of the \textit{update\_phi} kernel. The
naive and shared strategies are labeled NAIVE and SHARED respectively. Further,
each flavor of the code generation strategy is labeled by GEN followed by the 3
choices that identify it.
%
The wide performance range inhibits the readability of the figure,
therefore, a more focused Figure~\ref{titanx-w1-sweep}(b) is provided. As would
be expected, the naive strategy exhibits the worst performance over all
thread block
sizes, as it does not explicitly cache repeated device memory read operations.
The SHARED and \textit{GEN-SSS} strategies come next in terms of performance.
Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
differ in one aspect. Namely, \textit{GEN-SSS} explicitly unrolls the internal
loops of the kernel. However, there is no significant performance difference
between them. The other flavors of the code generation strategy attain higher
performance as they unroll internal loops as well as cache data in
registers. The \mbox{Titan-X} obtains the best performance with the \textit{GEN-RSS}
strategy and a thread block size of 64. The results of explicitly vectorized kernels
are omitted as they obtain worse performance on the \mbox{Titan-X}.

@@

\begin{comment}
% K40
\begin{figure*}[t]	% [hbt]
  \centering
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the K40c GPU, with
  explicit kernel vectorization of width 2, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{k40-w2-sweep}
\end{figure*}
\end{comment}

% \looseness=-1
Figures~\ref{k40-w2-sweep} (a) and (b) present the performance of the K40c GPU
for the same experimental configuration as shown for the \mbox{Titan-X} in
Figure~\ref{titanx-w1-sweep} with one exception. Namely, the code version used
for this plot is explicitly vectorized with a vector width of~2. The results
for the other code versions with vector width 4 and no vectorization are
omitted as they exhibit lower performance.
%
Surprisingly, Figure~\ref{k40-w2-sweep} shows that the NAIVE strategy
outperforms SHARED and some of the GEN strategies. This can be explained by the
unique properties of the Tesla Super Computing line of products to which the
K40c belongs. These GPUs include enhanced L2 caching mechanisms that
accelerate repeated and sparse memory accesses. This is especially advantageous
as it caches repeated reads across streaming multiprocessors. However, the
highest performance is attained by \textit{GEN-RRS} which explicitly employs
registers for both $Probs_i$ and $Grads_i$.

The performance results obtained from the GTX~\mbox{Titan-X} and Tesla~K40c GPUs
reinforce the importance of customizing compute kernels to each GPU's specific
architecture and capabilities. For instance, each GPU achieved its highest
performance by employing a different strategy. Moreover, each GPU displayed
different strategy-performance orderings.

@@

\subsubsection{Comparison of Compute Devices}
Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
GTX~\mbox{Titan-X}, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
sequential C++ version. These results are consistent with the relative
capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
the \mbox{Titan-X} achieves the highest speedup of 86 relative to the baseline. At the
other end, the parallel CPU version attains a speedup factor of~20.9.

As vectorization is a popular GPU optimization, it is important to note that
both GTX GPUs achieved their highest performance with non-vectorized kernels.
On the other hand, the Tesla GPUs obtained the best performance with a vector
width of~2. Figure~\ref{gpus-v-sweep} presents the execution time of the best
performing strategy for each of the 4 GPUs. In this figure, the performance is
normalized over the non-vectorized kernel version for each GPU. For instance,
the \mbox{Titan-X} exhibits an overhead factor of approximately 1.8 when using a vector
width of~4. On the other hand, the K40c improves its performance by roughly
20\% when it uses a vector width of~2 compared to the non-vectorized kernel.
Therefore, explicit vectorization of the kernels can be either useful or
harmful depending on the GPU architecture and the specific problem it is
applied to.

### 2017-05-10 gpu-05-05-perplexity.tex
Begin working on picture refactor

\begin{comment}
\begin{figure}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
  \caption{Perplexity convergence of com-DBLP on the GTX~Titan-X~GPU. Number
  of communities chosen to maximally fill the GPU memory.}
  \label{fig-ppx-gpu}
\end{figure}

@@

\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu-cpu.eps, width=\textwidth}
  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~Titan-X~CPU.
  (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
  maximally fill the memory of GPU and CPU respectively.}
  \label{fig-ppx-gpu-cpu}
\end{figure*}

@@

Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
com-DBLP, with $N=317,080$. It is computed on the GTX Titan-X GPU with input
parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.

Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity development convergence for
com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
$M$=4096, |\Neighbors|=32. Sufficient convergence is reached after 18~hours. Note
that this network graph is so large that the GTX Titan-X would be unable to
process more than $K$=256 communities.

>>>>

\begin{comment}
\begin{figure}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
  \caption{Perplexity convergence of com-DBLP on the GTX~\mbox{Titan-X}~GPU. Number
  of communities chosen to maximally fill the GPU memory.}
  \label{fig-ppx-gpu}
\end{figure}

@@

\begin{figure}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu-cpu.eps, width=0.45\textwidth}
  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~\mbox{Titan-X}~CPU.
  (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
  maximally fill the memory of GPU and CPU respectively.}
  \label{fig-ppx-gpu-cpu}
\end{figure}

@@

Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
com-DBLP, with $N=317,080$. It is computed on the GTX~\mbox{Titan-X} GPU with input
parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.

Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity convergence for
com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
$M$=4096, $|\Neighbors|$=32. Sufficient convergence is reached after 18~hours. Note
that this network graph is so large that the GTX~\mbox{Titan-X} would be unable to
process more than $K$=256 communities.

### 2017-05-10 gpu-05-evaluation.tex
Begin working on picture refactor

All experiments were conducted on the VU Amsterdam DAS5 cluster. The cluster
consists of 68 compute nodes each equipped with a dual 8-core Intel Xeon
E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and 8TB of storage.
Additionally, the cluster is fitted with a number of Nvidia GPUs including GTX
TitanX, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
their properties. Table~\ref{table-snap} outlines the graph
properties of the networks used to evaluate the algorithm's performance. These
networks were obtained from the SNAP collection~\cite{snapnets}.

\begin{table}[tb]
\center\begin{tabular}{l r r c}
Network & \#Vertices & \#Edges \\
\hline
CA-HepPh        &    12,008 &    118,521 \\
com-DBLP        &   317,080 &  1,049,866 \\
% com-Youtube     & 1,134,890 &  2,987,624 \\
com-LiveJournal & 3,997,962 & 34,681,189 \\
\hline
\end{tabular}
\caption{Network graphs from the Stanford Snap collection}
\label{table-snap}
\end{table}

\input{05-02-sequential}
\input{05-03-accelerator}
\input{05-05-perplexity}

>>>>

All experiments were conducted on the VU Amsterdam DAS5 cluster. The cluster
consists of 68 compute nodes each equipped with a dual 8-core Intel Xeon
E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and 8TB of storage.
Additionally, the cluster is fitted with a number of Nvidia GPUs including
GTX~\mbox{Titan-X}, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
their properties. Table~\ref{table-snap} outlines the graph
properties of the networks used to evaluate the algorithm's performance. These
networks were obtained from the SNAP collection~\cite{snapnets}.

\begin{comment}
\begin{table}[tb]
\center\begin{tabular}{l r r c}
Network & \#Vertices & \#Edges \\
\hline
CA-HepPh        &    12,008 &    118,521 \\
com-DBLP        &   317,080 &  1,049,866 \\
% com-Youtube     & 1,134,890 &  2,987,624 \\
com-LiveJournal & 3,997,962 & 34,681,189 \\
\hline
\\[-1ex]
\end{tabular}
\caption{Network graphs from the Stanford Snap collection}
\label{table-snap}
\end{table}
\end{comment}

\begin{table}
  \centering
  \def\tabcolsep{0.2em}
  \begin{tabular}{l r r r p{9em}}
    Name            & \#Vertices &       \#Edges & \multicolumn{1}{c}{\#Ground-} & Description \\
                    &            &               & \multicolumn{1}{c}{truths}    &             \\
    \hline
    CA-HepPh        &    12,008  &    118,521    & \multicolumn{1}{c}{n/a}       & Paper citation network \\
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
    \hline
    \\[-1ex]
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{table-snap}
\end{table}

\input{gpu-05-02-sequential}
\input{gpu-05-03-accelerator}
\input{gpu-05-05-perplexity}
